---
title: "<br> NYPD Shooting Incident Data Report <br><br>" 
author: "<font size =6, color='white'>Burak Kebapci</font>"
date: "<font size =6, color='white'>Nov 11th, 2021</font>"
---

```{=html}
<hr><p id="MainPoint"> Analysis and Interpretations </p><hr>
```
```{r}

NYPD = data.frame(read_csv("NYPD_Shooting_Incident_Data__Historic_.csv", skip_empty_rows = TRUE), stringsAsFactors = T)


### Removing unwanted variables/ columns
### 

NYPD <- NYPD %>%
    select(-INCIDENT_KEY, -X_COORD_CD, -Y_COORD_CD, -Latitude, -Longitude, -Lon_Lat)


NYPD[duplicated(NYPD$INCIDENT_KEY),]



Datatypelist<-as.data.frame(skimr::skim(NYPD)[[1]],skimr::skim(NYPD)[[2]])

knitr::kable(Datatypelist, "pipe", col.names = "Data Type")

#$`skimr::skim(NYPD)[[1]]`=="character"

```

```{r}
#char_vector <- c("OCCUR_DATE","BORO","LOCATION_DESC","PERP_AGE_GROUP","PERP_SEX","PERP_RACE","VIC_AGE_GROUP","VIC_SEX","VIC_RACE","Lon_Lat")

# Putting unknown for missing string values
NYPD[Datatypelist$`skimr::skim(NYPD)[[1]]`=="character"][is.na(NYPD[Datatypelist$`skimr::skim(NYPD)[[1]]`=="character"])] <- "UNKNOWN"


# Putting 0 for missing numeric values
NYPD[Datatypelist$`skimr::skim(NYPD)[[1]]`=="numeric"][is.na(NYPD[Datatypelist$`skimr::skim(NYPD)[[1]]`=="numeric"])] <- 0
```

#### We have taken care of missing values. Lets us begin with our analysis and understand the dataset. Here we go through each variable/feature column and get meaningful insights from them. As we proceed, we will keep updating our dataset so that by the end, we get a complete dataset with variables and values we need for our further analysis.

#### We can see above that the frequency of the Boro is plotted. For this dataset, we have five Boros. Among them, Brooklyn has the highest records, while Staten Island has the lowest record of shootouts. Can we say that Brooklyn is more dangerous than the rest of Boros? Let us find it out further in our report.

```{=html}
<hr><p id="MainPoint"> Our Methodology </p><hr>
```
### Here, we proceed with identifying the methods we will be using, along with justification for those methods.

# Business Questions

## 1. Is there a relationship between the Race of the victim and the Perpetrator who died in the shooting incident? *Major contribution by Chaya Kotturesh*

#### Here, we will use the Generalized regression model to solve this question. We know that Black people have a higher count of shootings when compared with others. Here are predictable variable is the Perpetrator's race, and the independent variable will be Victim race. We will create a model for those who are dead in the shootings.

```{r warning=FALSE, error=FALSE}
nypd_df <- read.csv("Data\\NYPD_Shooting_Incident_Data__Historic_.csv")
char_vector <- c("OCCUR_DATE","BORO","LOCATION_DESC","PERP_AGE_GROUP","PERP_SEX","PERP_RACE","VIC_AGE_GROUP","VIC_SEX","VIC_RACE","Lon_Lat")
nypd_df[char_vector][is.na(nypd_df[char_vector])] <- "UNKNOWN"


```

#### Reading the NYPD data set and performed the cleaning of character columns by filling UNKNOWN for NA's

```{r}
death_data<- nypd_df %>% filter(nypd_df$STATISTICAL_MURDER_FLAG== TRUE)

tab1(death_data$VIC_RACE, cex.main=1, cex.name=0.8, cex.axis=0.8, cex.lab=0.8, sort.group ="decreasing", bar.values ="percent", main = paste(GetFigureCount()," | Percentage of victim's died"), xlab=" % of shoot outs", ylab="Race of Victims", horiz = T)

```

#### As the business question is to find the relationship between the race of people that lead to the victim's death, I filtered the data set based on the victim's death. As we can see, there are around 4,127 records of shootings that lead to the victim's deaths. In that, the highest number of victims who died were of Black Race with whopping 70%, followed by White Hispanics with 15.9%

```{r}
death_data$PERP_RACE <- sub("^$", "UNKNOWN", death_data$PERP_RACE)#replacing empty cells with UNKNOWN
death_data$VIC_RACE <- sub("^$", "UNKNOWN", death_data$VIC_RACE)#replacing empty cells with UNKNOWN

crosstable2<- table(death_data$PERP_RACE,death_data$VIC_RACE)
crosstable2 <- round(prop.table(crosstable2)*100, 2)
crosstable2<-kable(crosstable2, caption = paste(GetFigureCount(), " | Percentage of Perpetrator's race killing victim's race"), xlab="test") %>% 
    kableExtra::kable_styling(., position = "float_left")
crosstable2

```

#### Intending to find the relationship between Perpetrator and victim's race, we created the contingency table.

#### We first convert the data as a table.

```{r fig.height = 7, fig.width = 18.5}
#library("gplots")
# 1. convert the data as a table
test<- table(death_data$PERP_RACE,
             death_data$VIC_RACE)
dt <- as.table(as.matrix(test))
```

#### Now, we use the converted data to visualize the Perpetrator's race killing the victim's race.

```{r fig.height = 7, fig.width = 18.5}
# 2. Graph
balloonplot(t(dt), 
            main =paste(GetFigureCount(), " | Perpetrator's race killing victim's race"), 
            xlab ="Victim's death", 
            ylab="Perpetrator's Race",
            label = FALSE, 
            show.margins = FALSE, 
            cex.main = 0.9  )
```

#### From the above graph, we can observe that Black and Unknown or white-Hispanic races murdered most Blacks. Unfortunately, the following significant pattern can be found where White-Hispanic were majorly murdered by White Hispanic themselves and next highest by BLACK people.

#### Even though contingency tables show some relationship between, Perpetrator and the victim's race, let us identify by Performing a Chi-square test at 0.05 significant level to identify the relationship between two categorical variables in the contingency table.

### `df=(r−1)(c−1)` degrees of freedom and `p = 0.05` *("Chi-Square Test of Independence in R - Easy Guides - Wiki - STHDA", 2020).*

#### Here,

#### r is the number of rows in the contingency table

#### c is the number of column in the contingency table

```{=html}
<B><FONT SIZE =3 color="#778899">
Step a: State the hypothesis and identify the claim</font></B>
<br> 
<B>Null hypothesis</B> 
$\it{H_{0}}$: There is no relationship exists between the race of perpetrator and victim.
<br>
<B>Alternative hypothesis</B> 
$\it{H_{1}}$: There is relationship between the race of perpetrator and victim.
 
<B><FONT SIZE =3 color="#778899">
Step b: Find the critical value</FONT>
</B>
```
```{r}
#here rows=6 columns=6 so df= 5*5

Critical_Val = qchisq(.05, df=25)

cat("The Critical value is ",Critical_Val)


```

```{=html}
<B><FONT SIZE =3 color="#778899">Step c: Performing $\chi^2$ Test or find the p-value</FONT></B>

```
```{r}
output_task4 <-chisq.test(dt)
output_task4
cat("Our Decision mustbe :", ifelse(output_task4$p.value < 0.05, "\nReject the null Hypothesis", "\nFail to reject the null hypothesis" ))

```

```{=html}
<B><FONT SIZE =3 color="#778899">Step d: Make the decision</FONT></B>
<BR>The decision rule is that if $p$-value is lesser than $\alpha$ we must reject the null hypothesis.
$p$-value < $\alpha$ i.e. 0.00000000000000022 is lesser than  0.05 ,so we must reject the $\it{H_{0}}$

<B><FONT SIZE =3 color="#778899">Step e: Summary</FONT></B>
<BR>
```
#### Interpretation: The evidence leads us to reject the null hypothesis. Therefore, there is a relationship between the race of Perpetrator and victim.

#### Another approach to figure out the relationship between categorical variables in R is by using Cramer's V coefficient

```{r fig.height = 15, fig.width = 18.5}
##library(vcd)

assocstats(table(death_data$PERP_RACE, death_data$VIC_RACE))

```

### interpretation of this function:

##### 1. Cramer's V varies from 0 to 1, a value of 1 indicates perfect association.In this scenario Cramer's V co-efficient is 0.334 which shows a moderate effect due to association.A value between `0.30 – < 0.50` is considered to be moderate.

#### 2. Contingency coefficient values also vary between 0 to 1. The larger the contingency coefficient, the stronger the association. 0.5 again shows the moderate association between Perpetrator and victim's race.

#### The discipline, the relevant data, and the analyst's goals all influence how measures of association are interpreted. There are sometimes guidelines for "small," "medium," and "large" impact. A smaller effect size may be considered "large," but in physical science, such as chemistry, it may be considered very small in psychology or behavioral science. The unique circumstances of the study are necessary as well.

```{=html}
<br><hr><br>
```